---
title: "Spike Test vs. Rolling Standard Deviation"
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
params: 
  county: county
  spike_table: spike_table
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(data.table)
library(DT)
library(dplyr)
library(ggplot2)
library(gtools)
library(here)
library(knitr)
library(lubridate)
library(plotly)
library(purrr)
library(qaqcmar)
library(readr)
library(strings)
library(sensorstrings)
library(summaryplots)
library(tidyr)
library(viridis)

source(here("functions/subchunkify.R"))

theme_set(theme_light())

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

county <- params$county

# reprocessed data (no preliminary QC)
dat_raw <- readRDS(here("data/temp_rolling_sd_reprocessed.rds")) %>%
  select(-c(int_sample, n_sample, n_sample_effective, sd_roll, rolling_sd_flag_temperature_degree_c)) %>%
  rename(temperature_degree_c = value_temperature_degree_c) %>% 
  filter(county == county)

```

```{r}
#| echo: false
#| message: false

# spike_table <- q_thresholds %>%
#   select(spike_low = q_999) %>%
#   mutate(
#     variable = "temperature_degree_c",
#     spike_high = 3 * spike_low
#   ) %>%
#   select(variable, spike_high, spike_low)

spike_table <- params$spike_table

rolling_sd_table <- thresholds %>% 
  filter(
    qc_test == "rolling_sd", 
    variable == "temperature_degree_c"
  ) %>% 
  pivot_wider(names_from = "threshold", values_from = "threshold_value") %>% 
  select(variable, county, rolling_sd_max)

counties <- unique(dat_raw$county)
dat_qc <- NULL

for(i in seq_along(counties)) {
  
  county_i <- counties[i]
  
  dat_qc[[i]] <- dat_raw %>% 
    filter(county == county_i) %>% 
    qc_test_spike(spike_table = spike_table) %>% 
    rename(
      temperature_degree_c = value_temperature_degree_c,
      spike_flag = spike_flag_temperature_degree_c) %>%
    qc_test_rolling_sd(
      rolling_sd_table = rolling_sd_table,
      join_column = "county"
    )
}

dat_qc <- dat_qc %>% 
  map_df(bind_rows) %>% 
  rename(spike_flag_temperature_degree_c = spike_flag) %>% 
  mutate(
    sensor_type = NA,
    spike_flag_temperature_degree_c = as.numeric(spike_flag_temperature_degree_c)
  ) %>% 
  select(
    c(1:8), sensor_type, sensor_depth_at_low_tide_m,
    value_temperature_degree_c, 
    rolling_sd_flag_temperature_degree_c,
    spike_flag_temperature_degree_c
  )

flag_colours <- c("chartreuse4", "#E6E1BC", "#EDA247", "#DB4325", "grey24")

```

January 4, 2024

- This file compares spike test results and rolling standard deviation test results.
- Spike test thresholds are q99.9 of the pooled data.


```{r}
both <- dat_qc %>% 
  filter(rolling_sd_flag_temperature_degree_c > 2 & spike_flag_temperature_degree_c > 2)

rolling_sd <- dat_qc %>% 
  filter(rolling_sd_flag_temperature_degree_c > 2 & spike_flag_temperature_degree_c <= 2)

spike <- dat_qc %>% 
  filter(rolling_sd_flag_temperature_degree_c <= 2 & spike_flag_temperature_degree_c > 2) %>% 
  mutate(spike_only = TRUE)

```

# Spike & Rolling SD Results

`r county`

Stations where there is at least one observation flagged by the spike test that is not flagged by the rolling sd test.

```{r, body} 
#| results: 'asis'
#| message: FALSE

#include number of obs

stations <- unique(spike$station)
 
for(i in seq_along(stations)){
  
  station_i <- stations[i]
  
  cat('\n##', station_i, '\n')
  
  cat('\n')
  
  # subset data to station of interest
  dat_i <- dat_qc %>%
    filter(station == station_i) %>%
    qc_pivot_longer(qc_tests = c("spike", "rolling_sd")) %>% 
    ss_convert_depth_to_ordered_factor()
  
  # height of figure j
  n_sensor <- length(unique(dat_i$sensor_serial_number))
  
  if(n_sensor == 1) h = 2.25
  if(n_sensor > 1) h = 2.25 * n_sensor
  
  p <- qc_plot_flags(dat_i, qc_tests = c("spike", "rolling_sd"))
 
  subchunkify(p, fig_height = h, fig_width = 8.5)
  
  cat('\n')
}

```

# Only Spike Test

Observations flagged by the spike test but not the rolling sd test.

## Number of Observations
```{r}

spike %>% 
  group_by(county, station) %>% 
  summarise(n = n()) %>% 
  datatable(options = dt_options, rownames = FALSE)

```

## Observations

```{r} 
#| results: 'asis'
#| message: FALSE

#include number of obs

stations <- unique(spike$station)
 
for(i in seq_along(stations)){
  
  station_i <- stations[i]
  
  cat('\n##', station_i, '\n')
  
  cat('\n')
  
  # subset data to station of interest
  dat_i <- dat_qc %>%
    filter(station == station_i) %>%
    qc_pivot_longer(qc_tests = c("spike", "rolling_sd")) %>% 
    ss_convert_depth_to_ordered_factor() %>% 
    left_join(
      spike %>% 
        filter(station == station_i) %>% 
        select(county, station, deployment_range, timestamp_utc,
               sensor_serial_number, spike_only), 
      by = c(
        "county", "station", "deployment_range", "timestamp_utc",
        "sensor_serial_number")
    ) %>% 
    mutate(spike_only = if_else(is.na(spike_only), FALSE, spike_only))
  
  # height of figure j
  n_sensor <- length(unique(dat_i$sensor_serial_number))
  
  if(n_sensor == 1) h = 2.5
  if(n_sensor > 1) h = 2.5 * n_sensor
  
  p <- ggplot(dat_i, aes(timestamp_utc, value, col = spike_only)) +
    geom_point() +
    scale_color_manual(values = c(flag_colours[1], flag_colours[3]), guide = "none") +
    facet_wrap(~sensor_serial_number, ncol = 1)
 
  subchunkify(p, fig_height = h, fig_width = 8.5)
  
  cat('\n')
}

```
