---
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8, fig.height = 6)

library(dplyr)
library(DT)
library(here)
library(ggplot2)
library(leaflet)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(readr)
library(strings)
library(summaryplots)
library(tidyr)
library(viridis)

source(here("functions/remove_do_correction.R"))
source(here("functions/subchunkify.R"))

theme_set(theme_light())

x_axis <- scale_x_continuous("Dissolved Oxygen (mg/L)")

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

# data - preliminary QC 
dat_all <- import_strings_data(input_path = here("data-raw")) %>% 
  remove_do_correction(keep_temp = TRUE) %>% 
  select(COUNTY, WATERBODY, STATION, 
         DEPLOYMENT_PERIOD,
         TIMESTAMP, DEPTH, VARIABLE, VALUE, UNITS) %>%
  #filter(VARIABLE == "Dissolved Oxygen", UNITS == "mg/L") %>%
  mutate(
    DEPTH = factor(round(as.numeric(DEPTH))),
    MONTH = month(TIMESTAMP),
    YEAR = factor(year(TIMESTAMP))
  ) 

county_pal <- get_county_colour_palette(length(unique(dat_all$COUNTY)))
depth_pal <- viridis(length(unique(dat_all$DEPTH)), direction = -1)
year_pal <- viridis( length(unique(dat_all$YEAR)), option = "C")

theme_facet_plotly <- theme(
  panel.spacing.y = unit(30, "lines"),
  panel.spacing.x = unit(20, "lines")
)

```

# Dissolved Oxygen Distribution

August 15, 2023

- Data submitted to the Open Data Portal in December 2022.

- Preliminary QC (submitted to Open Data Portal)
  - Obvious outliers removed.
  - Suspected biofouling removed.
  
- Notes on salinity correction:
  - The DO concentration data sent to the Open Data Portal in December 2022 was corrected for salinity using `strings::do_salinity_correction()`.
  - $DO_{corrected} = DO_{raw} * F_s$, where $F_s$ is a salinity correction factor between 0 and 1.
  - The newly processed and flagged data will NOT be corrected for salinity.
  - So to calculate thresholds based on historical data, the correction factor must be removed before analysis.
  - The data presented here is the *uncorrected* dissolved oxygen data.

## Deployment Dates

### Table 1
```{r}
dat_all %>% 
  distinct(COUNTY, STATION, DEPLOYMENT_PERIOD) %>% 
  arrange(COUNTY, STATION) %>% 
  select(-COUNTY) %>% 
  datatable(options = dt_options, rownames = FALSE)

```

## Preliminary QC

## Halifax

### Shut-In Island
```{r}
dat_all %>% 
  filter(STATION == "Shut-In Island") %>% 
  ggplot_variables_at_depth()
```

## Lunenburg

### Birchy Head
```{r}
dat_all %>% 
  filter(STATION == "Birchy Head") %>% 
  ggplot_variables_at_depth()
```

### Flat Island
```{r}
dat_all %>% 
  filter(STATION == "Flat Island") %>% 
  ggplot_variables_at_depth()
```

### Little Rafuse Island
```{r}
dat_all %>% 
  filter(STATION == "Little Rafuse Island") %>% 
  ggplot_variables_at_depth()
```

### Tilley Point
```{r}
dat_all %>% 
  filter(STATION == "Tilley Point") %>% 
  ggplot_variables_at_depth()
```

### Upper Blanford
```{r}
dat_all %>% 
  filter(STATION == "Upper Blandford") %>% 
  ggplot_variables_at_depth()
```

# QC Recommendations

- Shut-In Island: ~ Aug 15. DO does not follow temperature trend.
- Birchy Head: ~ Aug 15. DO does not follow temperature trend.
- Flat Island: July 1. DO trends down most of the deployment, likely due to sensor drift.
- Little Rafuse Island: June 10. DO trends down most of the deployment, likely due to sensor drift.
- Tilley Point: leave as-is; DO trend loosely corresponds to spikes in temperature.
- Upper Blandford: ~ Aug 15. DO does not follow temperature trend.

```{r}
dat_filt <- dat_all %>% 
    filter(
      !(STATION %in% c("Shut-In Island", "Birchy Head", "Upper Blandford") & 
      #  UNITS == "mg/L" & 
        TIMESTAMP > as_datetime("2022-08-15")),
    !(STATION == "Flat Island"  & 
      #  UNITS == "mg/L" &
        TIMESTAMP > as_datetime("2022-07-01")), 
    !(STATION == "Little Rafuse Island" &  
       # UNITS == "mg/L" &
        TIMESTAMP > as_datetime("2022-06-10")) 
  #  UNITS != "degrees Celsius"
    )
  

# dat_filt2 <- dat_all %>% 
#   filter(
#     (STATION == "Shut-In Island" & TIMESTAMP <= as_datetime("2022-08-15")) |
#       (STATION == "Birchy Head" & TIMESTAMP <= as_datetime("2022-08-15")) |
#       (STATION == "Flat Island" & TIMESTAMP <= as_datetime("2022-07-01")) |
#       (STATION == "Little Rafuse Island" &
#          TIMESTAMP <= as_datetime("2022-06-10")) |
#       (STATION == "Upper Blandford" & TIMESTAMP <= as_datetime("2022-08-15")) |
#     STATION == "Tilley Point",
#     UNITS == "mg/L"
#   )

```

```{r}
dat_all_cor <- dat_all %>% 
  select(-UNITS) %>% 
  pivot_wider(names_from = "VARIABLE", values_from = "VALUE") %>% 
  group_by(STATION) %>% 
  summarise(correlation = cor(`Dissolved Oxygen`, Temperature))

dat_filt_cor <- dat_filt %>% 
  select(-UNITS) %>% 
  pivot_wider(names_from = "VARIABLE", values_from = "VALUE") %>% 
  group_by(STATION) %>% 
  summarise(correlation_filt = cor(`Dissolved Oxygen`, Temperature))

corr <- dat_all_cor %>% 
  left_join(dat_filt_cor)

```


# Additional QC

- Note difference in scales (x-axis, y-axis, colour) compared to previous figures.

## Halifax

### Shut-In Island
```{r}
dat_filt %>% 
  filter(STATION == "Shut-In Island") %>% 
  ggplot_variables_at_depth(color.pal = "#007500") 
```

## Lunenburg

### Birchy Head
```{r}
dat_filt %>% 
  filter(STATION == "Birchy Head") %>% 
  ggplot_variables_at_depth(color.pal = "#007500") 
```

### Flat Island
```{r}
dat_filt %>% 
  filter(STATION == "Flat Island") %>% 
  ggplot_variables_at_depth(color.pal = "#007500") 
```

### Little Rafuse Island
```{r}
dat_filt %>% 
  filter(STATION == "Little Rafuse Island") %>% 
  ggplot_variables_at_depth(color.pal = "#007500")
```

### Tilley Point
```{r}
dat_filt %>% 
  filter(STATION == "Tilley Point") %>% 
  ggplot_variables_at_depth(color.pal = "#007500") 
```

### Upper Blanford
```{r}
dat_filt %>% 
  filter(STATION == "Upper Blandford") %>% 
  ggplot_variables_at_depth(color.pal = "#007500") 
```


