---
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(DT)
library(here)
library(ggplot2)
library(leaflet)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(readr)
library(strings)
library(summaryplots)
library(tidyr)
library(viridis)

source(here("functions/remove_do_correction.R"))
source(here("functions/subchunkify.R"))

theme_set(theme_light())

x_axis <- scale_x_continuous("Dissolved Oxygen (mg/L)")

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

# data - preliminary QC 
dat_all <- import_strings_data(input_path = here("data-raw")) %>%
  remove_do_correction() %>% 
  select(COUNTY, WATERBODY, STATION, TIMESTAMP, DEPTH, VARIABLE, VALUE, UNITS) %>%
  mutate(
    DEPTH = factor(round(as.numeric(DEPTH))),
    MONTH = month(TIMESTAMP),
    YEAR = factor(year(TIMESTAMP))
  ) 

# filtered data
dat_filt <- dat_all %>% 
  filter(
    !(STATION %in% c("Shut-In Island", "Birchy Head", "Upper Blandford") & 
        TIMESTAMP > as_datetime("2022-08-15")),
    !(STATION == "Flat Island"  &  TIMESTAMP > as_datetime("2022-07-01")), 
    !(STATION == "Little Rafuse Island" &  TIMESTAMP > as_datetime("2022-06-10")) 
  )

county_pal <- get_county_colour_palette(length(unique(dat_all$COUNTY)))
depth_pal <- viridis(length(unique(dat_all$DEPTH)), direction = -1)
year_pal <- viridis( length(unique(dat_all$YEAR)), option = "C")

theme_facet_plotly <- theme(
  panel.spacing.y = unit(30, "lines"),
  panel.spacing.x = unit(20, "lines")
)

```

# Dissolved Oxygen Distribution

August 15, 2023

- Data submitted to the Open Data Portal in December 2022.

- Preliminary QC (submitted to Open Data Portal)
  - Obvious outliers removed.
  - Suspected biofouling removed.

- Filtered Data
  - Additional suspected biofouling was removed from the May 2022 deployments.
  
- Notes on salinity correction:
  - The DO concentration data sent to the Open Data Portal in December 2022 was corrected for salinity using `strings::do_salinity_correction()`.
  - $DO_{corrected} = DO_{raw} * F_s$, where $F_s$ is a salinity correction factor between 0 and 1.
  - The newly processed and flagged data will NOT be corrected for salinity.
  - So to calculate thresholds based on historical data, the correction factor must be removed before analysis.
  - The data presented here is the *uncorrected* dissolved oxygen data.

## Preliminary QC 

### Pooled

::: panel-tabset

#### Figure 1: All observations
```{r}
#| label: fig-dist-all
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L).

p <- plot_histogram(dat_all, hist_col = VALUE, binwidth = 0.5) +
  x_axis

ggplotly(p)
```

#### Figure 2: By depth
```{r}
#| label: fig-dist-all-depth
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5,
  fill_col = DEPTH, pal = depth_pal
) +
  x_axis

ggplotly(p)
```

#### Figure 3: By year
```{r}
#| label: fig-dist-all-year
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, 
  fill_col = YEAR, pal = year_pal
) +
  x_axis

ggplotly(p)
```

:::

### County

::: panel-tabset

#### Figure 4: County 
```{r}
#| label: fig-dist-county
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations by county (binwidth = 0.5 mg/L).

p <- dat_all %>% 
  plot_histogram(hist_col = VALUE, binwidth = 0.5) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

#### Figure 5: County + Depth
```{r}
#| label: fig-dist-county-depth
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), within each county, coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = DEPTH, pal = depth_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly
  
ggplotly(p)
```

#### Figure 6: County + Year
```{r}
#| label: fig-dist-county-year
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), within each county coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```
:::

#### County + Month + Year

::: panel-tabset
```{r}
#| fig-height: 10
#| results: "asis"

counties <- unique(dat_all$COUNTY)

dat_all <- dat_all %>% 
  mutate(MONTH = month(MONTH, label = TRUE))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- dat_all %>% 
    filter(COUNTY == county_i) 
  
  cat('\n###', county_i, '\n')
  
  p <- plot_histogram(
    dat_i, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
  ) +
    x_axis +
    facet_wrap(~MONTH, ncol = 3)  
  
  print(p)
}

```
:::

## Deployments

The following figures show the deployments from these counties, with preliminary QC applied. See 0_do_mg_per_l_additional_qc.html for more context.

::: panel-tabset

```{r, body} 
#| results: 'asis'
#| message: FALSE

stations <- unique(dat_all$STATION)

for(i in seq_along(stations)){

  station_i <- stations[i]

  # subset data to station of interest
  dat_i <- dat_all %>%
    filter(STATION == station_i) %>% 
    convert_depth_to_ordered_factor()
  
  cat('\n##', unique(dat_i$COUNTY), '\n')
  
  cat('\n')
  
  p <- ggplot_variables_at_depth(dat_i) +
    scale_y_continuous("Dissolved Oxygen (mg/L)", limits = c(0, 18)) +
    theme(strip.text = element_blank())
  
  cat('\n###', station_i, '\n')
  
  subchunkify(p, fig_height = 3, fig_width = 8.5)
  
  cat('\n')
}
```
:::

## Filtered Data

### Pooled

::: panel-tabset

#### Figure 7: All observations
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L).

p <- plot_histogram(dat_filt, hist_col = VALUE, binwidth = 0.5) +
  x_axis

ggplotly(p)
```

#### Figure 8: By depth
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), coloured by depth.

p <- plot_histogram(
  dat_filt, hist_col = VALUE, binwidth = 0.5,
  fill_col = DEPTH, pal = depth_pal
) +
  x_axis

ggplotly(p)
```

#### Figure 9: By year
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, 
  fill_col = YEAR, pal = year_pal
) +
  x_axis

ggplotly(p)
```

:::

### County

::: panel-tabset

#### Figure 10: County 
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations by county (binwidth = 0.5 mg/L).

p <- dat_filt%>% 
  plot_histogram(hist_col = VALUE, binwidth = 0.5) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

#### Figure 11: County + Depth
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), within each county, coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = DEPTH, pal = depth_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly
  
ggplotly(p)
```

#### Figure 12: County + Year
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.5 mg/L), within each county coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```
:::

#### County + Month + Year

::: panel-tabset
```{r}
#| fig-height: 10
#| results: "asis"

counties <- unique(dat_all$COUNTY)

dat_filt<- dat_filt%>% 
  mutate(MONTH = month(MONTH, label = TRUE))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- dat_filt%>% 
    filter(COUNTY == county_i) 
  
  cat('\n###', county_i, '\n')
  
  p <- plot_histogram(
    dat_i, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
  ) +
    x_axis +
    facet_wrap(~MONTH, ncol = 3)  
  
  print(p)
}

```
:::



# Recommendations

- Histogram of Preliminary QC pooled data is left-skewed.
- Filtered data is still slightly left-skewed, mainly because of the Tilley Point 2022 deployment. This was not trimmed becuase the changes in DO loosely followed the inverse pattern of temperature.




