---
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(DT)
library(here)
library(ggplot2)
library(leaflet)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(readr)
library(sensorstrings)
library(strings)
library(summaryplots)
library(tidyr)
library(viridis)

source(here("functions/subchunkify.R"))

theme_set(theme_light())

x_axis <- scale_x_continuous("Dissolved Oxygen (mg/L)")

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

# data - preliminary QC 
dat_all <- readRDS(here("data/do_mg_per_l_rolling_sd_prelim_qc.rds")) %>% 
  select(
    -c(sensor_type, int_sample, n_sample, 
       rolling_sd_flag_dissolved_oxygen_uncorrected_mg_per_l)) %>% 
  mutate(
    depth = factor(sensor_depth_at_low_tide_m),
    year = factor(year(timestamp_utc)),
    month = month(timestamp_utc)
  ) %>% 
  na.omit() 

# filtered data
dat_filt <- dat_all %>% 
  filter(
    !(station %in% c("Shut-In Island", "Birchy Head", "Upper Blandford") & 
        timestamp_utc > as_datetime("2022-08-15")),
    !(station == "Flat Island"  &  timestamp_utc > as_datetime("2022-07-01")), 
    !(station == "Little Rafuse Island" &  timestamp_utc > as_datetime("2022-06-10")) 
  )

county_pal <- get_county_colour_palette(length(unique(dat_all$county)))
depth_pal <- viridis(length(unique(dat_all$sensor_depth_at_low_tide_m)), direction = -1)
year_pal <- viridis( length(unique(dat_all$year)), option = "C")

theme_facet_plotly <- theme(
  panel.spacing.y = unit(30, "lines"),
  panel.spacing.x = unit(20, "lines")
)

```

# Dissolved Oxygen (mg/L): Rolling Standard Deviation Distribution

DRAFT August 16, 2023

- Data submitted to the Open Data Portal in December 2022.
- Depth rounded to nearest whole number

- Rolling standard deviation is not normally distributed.
- Distribution is skewed right.
- For thresholds, we are concerned about high rolling standard deviations.
  - Eventually could use rolling standard deviation = 0 as a type of flat line test.
- Consider using upper quartiles (e.g., 95, 97, 99, 99.7) as threshold values instead of mean + 3 * sd.


## Preliminary QC 

### Pooled

::: panel-tabset

#### Figure 1: All observations
```{r}
#| label: fig-dist-all
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L).

p <- plot_histogram(dat_all, hist_col = sd_roll, binwidth = 0.25) +
  x_axis

ggplotly(p)
```

#### Figure 2: By depth
```{r}
#| label: fig-dist-all-depth
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = sd_roll, binwidth = 0.25,
  fill_col = depth, pal = depth_pal
) +
  x_axis

ggplotly(p)
```

#### Figure 3: By year
```{r}
#| label: fig-dist-all-year
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), coloured by year.

p <- plot_histogram(
  dat_all, hist_col = sd_roll, binwidth = 0.25, 
  fill_col = year, pal = year_pal
) +
  x_axis

ggplotly(p)
```

:::

### County

::: panel-tabset

#### Figure 4: County 
```{r}
#| label: fig-dist-county
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations by county (binwidth = 0.25 mg/L).

p <- dat_all %>% 
  plot_histogram(hist_col = sd_roll, binwidth = 0.25) +
  x_axis +
  facet_wrap(~county, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

#### Figure 5: County + Depth
```{r}
#| label: fig-dist-county-depth
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), within each county, coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = sd_roll, binwidth = 0.25, fill_col = depth, pal = depth_pal
) +
  x_axis +
  facet_wrap(~county, scales = "free", ncol = 3) +
  theme_facet_plotly
  
ggplotly(p)
```

#### Figure 6: County + Year
```{r}
#| label: fig-dist-county-year
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), within each county coloured by year.

p <- plot_histogram(
  dat_all, hist_col = sd_roll, binwidth = 0.25, fill_col = year, pal = year_pal
) +
  x_axis +
  facet_wrap(~county, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```
:::

#### County + Month + Year

::: panel-tabset
```{r}
#| fig-height: 12
#| results: "asis"

counties <- unique(dat_all$county)

dat_all <- dat_all %>% 
  mutate(month = month(month, label = TRUE))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- dat_all %>% 
    filter(county == county_i) 
  
  cat('\n###', county_i, '\n')
  
  p <- plot_histogram(
    dat_i, hist_col = sd_roll, binwidth = 0.25, fill_col = year, pal = year_pal
  ) +
    x_axis +
    facet_wrap(~month, ncol = 3)  
  
  print(p)
}

```
:::


### Summary Statistics

::: panel-tabset

#### Figure 7: Central Tendancy
```{r}
#| fig-height: 6
#| fig-cap: The mean and standard deviation of rolling standard deviation in each county.
#| label: fig-mean-filt

stats_all <- dat_all %>% 
  summarise(
    max = max(sd_roll),
    mean = mean(sd_roll),
    med = median(sd_roll),
    stdev = sd(sd_roll),
    
    q_90 = quantile(sd_roll, prob = 0.90),
    q_95 = quantile(sd_roll, prob = 0.95),
    q_97 = quantile(sd_roll, prob = 0.97),
    q_99 = quantile(sd_roll, prob = 0.99),
    q_997 = quantile(sd_roll, prob = 0.997)
  ) %>% 
  mutate(county = "All") %>% 
  bind_rows(
    dat_all %>% 
      group_by(county) %>% 
      summarise(
        max = max(sd_roll),
        mean = mean(sd_roll),
        med = median(sd_roll),
        stdev = sd(sd_roll),
        
        q_90 = quantile(sd_roll, prob = 0.90),
        q_95 = quantile(sd_roll, prob = 0.95),
        q_97 = quantile(sd_roll, prob = 0.97),
        q_99 = quantile(sd_roll, prob = 0.99),
        q_997 = quantile(sd_roll, prob = 0.997)
      ) 
  ) %>%
  mutate(
    variable = "dissolved oxygen_percent_saturation",
    thresh_3sd = mean + 3 * stdev,
    across(.cols = where(is.numeric), .fns = round, digits = 2)
  ) %>% 
  select(county, everything())

p <- plot_mean_sd_county(stats_all)

ggplotly(p)
```

#### Figure 8: Quantiles
```{r}
#| fig-cap: Upper quantiles.
#| fig-height: 6
#| label: fig-quantiles-filt

p <- stats_all %>% 
  select(county, contains("q")) %>% 
  pivot_longer(cols = contains("q"), names_to = "quantile") %>% 
  ggplot(aes(quantile, value, group = county, col = county)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("grey20", county_pal), drop = FALSE) +
  scale_y_continuous(limits = c(0, 3))

ggplotly(p)
```

#### Table 1: Summary statistics
```{r}
stats_all %>% 
  select(-c(variable, max, contains("thresh"))) %>% 
  datatable(
    options = list(
      dom = 'ft',
      paging = FALSE,
      searching = TRUE,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    ), rownames = FALSE)
```

:::

## Filtered

### Pooled

::: panel-tabset

#### Figure 1: All observations
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L).

p <- plot_histogram(dat_filt, hist_col = sd_roll, binwidth = 0.25) +
  x_axis

ggplotly(p)
```

#### Figure 2: By depth
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), coloured by depth.

p <- plot_histogram(
  dat_filt, hist_col = sd_roll, binwidth = 0.25,
  fill_col = depth, pal = depth_pal
) +
  x_axis

ggplotly(p)
```

#### Figure 3: By year
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), coloured by year.

p <- plot_histogram(
  dat_filt, hist_col = sd_roll, binwidth = 0.25, 
  fill_col = year, pal = year_pal
) +
  x_axis

ggplotly(p)
```

:::

### County

::: panel-tabset

#### Figure 4: County 
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations by county (binwidth = 0.25 mg/L).

p <- dat_filt %>% 
  plot_histogram(hist_col = sd_roll, binwidth = 0.25) +
  x_axis +
  facet_wrap(~county, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

#### Figure 5: County + Depth
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), within each county, coloured by depth.

p <- plot_histogram(
  dat_filt, hist_col = sd_roll, binwidth = 0.25, fill_col = depth, pal = depth_pal
) +
  x_axis +
  facet_wrap(~county, scales = "free", ncol = 3) +
  theme_facet_plotly
  
ggplotly(p)
```

#### Figure 6: County + Year
```{r}
#| fig-height: 6
#| fig-cap: Distribution of dissolved oxygen observations (binwidth = 0.25 mg/L), within each county coloured by year.

p <- plot_histogram(
  dat_filt, hist_col = sd_roll, binwidth = 0.25, fill_col = year, pal = year_pal
) +
  x_axis +
  facet_wrap(~county, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```
:::

#### County + Month + Year

::: panel-tabset
```{r}
#| fig-height: 12
#| results: "asis"

counties <- unique(dat_filt$county)

dat_filt <- dat_filt %>% 
  mutate(month = month(month, label = TRUE))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- dat_filt %>% 
    filter(county == county_i) 
  
  cat('\n###', county_i, '\n')
  
  p <- plot_histogram(
    dat_i, hist_col = sd_roll, binwidth = 0.25, fill_col = year, pal = year_pal
  ) +
    x_axis +
    facet_wrap(~month, ncol = 3)  
  
  print(p)
}

```
:::


### Summary Statistics

::: panel-tabset

#### Figure 7: Central Tendancy
```{r}
#| fig-height: 6
#| fig-cap: The mean and standard deviation of rolling standard deviation in each county.

stats_all <- dat_filt %>% 
  summarise(
    max = max(sd_roll),
    mean = mean(sd_roll),
    med = median(sd_roll),
    stdev = sd(sd_roll),
    
    q_90 = quantile(sd_roll, prob = 0.90),
    q_95 = quantile(sd_roll, prob = 0.95),
    q_97 = quantile(sd_roll, prob = 0.97),
    q_99 = quantile(sd_roll, prob = 0.99),
    q_997 = quantile(sd_roll, prob = 0.997)
  ) %>% 
  mutate(county = "All") %>% 
  bind_rows(
    dat_filt %>% 
      group_by(county) %>% 
      summarise(
        max = max(sd_roll),
        mean = mean(sd_roll),
        med = median(sd_roll),
        stdev = sd(sd_roll),
        
        q_90 = quantile(sd_roll, prob = 0.90),
        q_95 = quantile(sd_roll, prob = 0.95),
        q_97 = quantile(sd_roll, prob = 0.97),
        q_99 = quantile(sd_roll, prob = 0.99),
        q_997 = quantile(sd_roll, prob = 0.997)
      ) 
  ) %>%
  mutate(
    variable = "dissolved oxygen_percent_saturation",
    thresh_3sd = mean + 3 * stdev,
    across(.cols = where(is.numeric), .fns = round, digits = 2)
  ) %>% 
  select(county, everything())

p <- plot_mean_sd_county(stats_all)

ggplotly(p)
```

#### Figure 8: Quantiles
```{r}
#| fig-cap: Upper quantiles.
#| fig-height: 6

p <- stats_all %>% 
  select(county, contains("q")) %>% 
  pivot_longer(cols = contains("q"), names_to = "quantile") %>% 
  ggplot(aes(quantile, value, group = county, col = county)) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("grey20", county_pal), drop = FALSE) +
  scale_y_continuous(limits = c(0, 3))

ggplotly(p)
```

#### Table 1: Summary statistics
```{r}
stats_all %>% 
  select(-c(variable, max, contains("thresh"))) %>% 
  datatable(
    options = list(
      dom = 'ft',
      paging = FALSE,
      searching = TRUE,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    ), rownames = FALSE)
```

:::
