---
title: "Dissolved Oxygen: Rolling Standard Deviation Thresholds Comparison"
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(data.table)
library(DT)
library(dplyr)
library(ggplot2)
library(gtools)
library(here)
library(knitr)
library(lubridate)
library(plotly)
library(qaqcmar)
library(strings)
library(sensorstrings)
library(summaryplots)
library(tidyr)
library(viridis)

source(here("functions/subchunkify.R"))

theme_set(theme_light())

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

dat_roll <- readRDS(here("data/do_mg_per_l_rolling_sd_reprocessed.rds")) %>%
  select(
    -c(int_sample, n_sample,
       rolling_sd_flag_dissolved_oxygen_uncorrected_mg_per_l)
  ) %>%
  na.omit() %>%
  separate(col = deployment_range, into = c("depl_date", NA), sep = " to ")

# do thresholds - exported from 3.1_do_rolling_sd_distribution
thresholds <- data.frame(0.51, 0.71,	0.86,	1.11, 1.38)
names(thresholds) <- c("q_90", "q_95", "q_97", "q_99", "q_997")

dat_summary <- dat_roll %>%
  mutate(
    thresh_q90 = case_when(
      sd_roll > thresholds$q_90 ~ 3,
      sd_roll <= thresholds$q_90 ~ 1,
      is.na(sd_roll) ~ 2),
    thresh_q95 = case_when(
      sd_roll > thresholds$q_95 ~ 3,
      sd_roll <= thresholds$q_95 ~ 1,
      is.na(sd_roll) ~ 2),
    thresh_q97 = case_when(
      sd_roll > thresholds$q_97 ~ 3,
      sd_roll <= thresholds$q_97 ~ 1,
      is.na(sd_roll) ~ 2),
    thresh_q99 = case_when(
      sd_roll > thresholds$q_99 ~ 3,
      sd_roll <= thresholds$q_99 ~ 1,
      is.na(sd_roll) ~ 2),
    thresh_q997 = case_when(
      sd_roll > thresholds$q_997 ~ 3,
      sd_roll <= thresholds$q_997 ~ 1,
      is.na(sd_roll) ~ 2)
  ) %>%
  pivot_longer(
    cols = contains("thresh"),
    names_to = "threshold",
    values_to = "threshold_value", names_prefix = "thresh_"
  ) %>%
  filter(threshold %in% c("q90", "q95", "q997"))

county_pal <- get_county_colour_palette(length(unique(dat_roll$county)))
flag_colours <- c("chartreuse4", "#E6E1BC", "#EDA247", "#DB4325", "grey24")

theme_facet_plotly <- theme(
  panel.spacing.y = unit(20, "lines"),
  panel.spacing.x = unit(15, "lines")
)
```

August 21, 2023

- This file compares potential rolling standard deviation thresholds.

- Potential thresholds are the upper quantiles (q90, q95, and q99.7) of the pooled data.

- Thresholds were calculated based on the filtered data. 
    - Suspected biofouling and obvious outliers were removed.
    - Fresh water and other outlier stations were removed.
    
- Additional suspected biofouling was removed from the May 2022 deployments.

- These thresholds are applied to the re-processed data to assess which threshold will be the most useful for automated QC.
  - No preliminary QC was applied to the reprocessed data.
  - A substantial number of these observations is expected to be biofouling.
  
- Potential thresholds:
```{r}
thresholds %>% 
  select(q_90, q_95, q_997) %>% 
  t() %>% 
  data.frame() %>% 
  rename(threshold_value = ".") %>% 
  datatable(
    width = "500px",
    list(
      dom = 'ft',
      searching = FALSE,
      paging = FALSE,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    )
  )
```


## Pooled

The percent of all observations flagged with each potential threshold.

::: panel-tabset

### Figure 1
```{r}
summary_all <- dat_summary %>%
  group_by(threshold) %>%
  summarise(n_obs = n(), n_suspect = sum(threshold_value == 3)) %>%
  mutate(percent_suspect = round((100 * n_suspect / n_obs), digits = 3)) %>%
  ungroup()

p <- ggplot(summary_all, aes(percent_suspect, threshold)) +
  geom_col(
    col = 1, fill = "#AE1759FF",
    position = position_dodge2(preserve = "single", padding = 0)
  ) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(
    "Percent of Observations", expand = expansion(mult = c(0, 0.1))
  ) +
  scale_y_discrete(name = "Threshold", limits = rev)

ggplotly(p)
```

### Table 1
```{r}
datatable(
  summary_all,
  options = list(
    dom = 'ft',
    paging = FALSE,
    searching = FALSE,
    columnDefs = list(list(className = 'dt-center', targets = "_all"))), rownames = FALSE)
```
:::

## By County

The percent of observations within each county flagged with each potential threshold.

::: panel-tabset

### Figure 2
```{r}
#| fig-height: 6
#| message: false

summary_county <- dat_summary %>%
  group_by(county, threshold) %>%
  summarise(n_obs = n(), n_suspect = sum(threshold_value == 3)) %>%
  mutate(percent_suspect = round((100 * n_suspect / n_obs), digits = 3)) %>%
  ungroup()

p <- ggplot(summary_county, aes(percent_suspect, threshold)) +
  geom_col(
    col = 1, fill = "#AE1759FF",
    position = position_dodge2(preserve = "single", padding = 0)
  ) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(
    "Percent of Observations", expand = expansion(mult = c(0, 0.1))
  ) +
  scale_y_discrete(name = "Threshold", limits = rev) +
  facet_wrap(~county, ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

### Table 2
```{r}
summary_county %>%
  select(-c(n_obs, n_suspect)) %>%
  pivot_wider(names_from = "threshold", values_from = "percent_suspect") %>%
  datatable(options = dt_options, rownames = FALSE)
```

:::

## By Deployment

The percent of observations within each deployment flagged with each potential threshold.

The `diff` column shows difference between the percent of observations flagged with the q_90 threshold and the q_997 threshold.

### Table 3
```{r}
#| message: false

summary_depl <- dat_summary %>%
  group_by(county, station, depl_date, threshold) %>%
  summarise(n_obs = n(), n_suspect = sum(threshold_value == 3)) %>%
  mutate(percent_suspect = round((100 * n_suspect / n_obs), digits = 3)) %>%
  ungroup() %>% 
  select(-c(n_obs, n_suspect)) %>% 
  pivot_wider(names_from = "threshold", values_from = "percent_suspect") %>%
  mutate(
    diff = q90 - q997,
    across(where(is.numeric), .fn = round, digits = 2)
  )

summary_depl %>% 
   datatable(options = dt_options, rownames = FALSE)
```

</br>

## Deployments with Largest `diff`

These figures show the flagged observations for each threshold for the deployments with the largest difference between thresholds.

::: panel-tabset

```{r}
#| results: "asis"

depl <- summary_depl %>% 
  arrange(desc(diff)) %>% 
  filter(diff >= 40)

thresh_plot <- unique(dat_summary$threshold)

for(i in 1:nrow(depl)){
 
  depl_i <- depl[i, 1:3]
  
  cat('\n###', paste(depl_i$station, depl_i$depl_date), '\n')
  
  for(j in seq_along(thresh_plot)) {
    
    cat('\n####', thresh_plot[j], '\n')
    
    dat_j <-  dat_summary %>% 
      filter(
        station == depl_i$station, 
        depl_date == depl_i$depl_date, 
        threshold == thresh_plot[j],
        value_dissolved_oxygen_uncorrected_mg_per_l > 0
        )  %>%
      mutate(
        depth = paste0(
          sensor_depth_at_low_tide_m, " m", 
          " (", sensor_serial_number, ")"),
        depth = ordered(depth, levels = gtools::mixedsort(unique(depth)))
      ) %>% 
      rename(sd_rolling_flag_value = threshold_value) %>% 
      qc_assign_flag_labels() 
    
    # height of figure j
    n_sensor <- length(unique(dat_j$sensor_serial_number))
    
    if(n_sensor == 1) h = 2.75
    if(n_sensor > 1) h = 2.75 * n_sensor
    
    p <- ggplot(
      dat_j, 
      aes(timestamp_utc, value_dissolved_oxygen_uncorrected_mg_per_l, 
          colour = sd_rolling_flag_value )
      ) +
      geom_point()+
      scale_y_continuous("Dissolved Oxygen (mg/L)") +
      scale_x_datetime("Date") +
      scale_colour_manual("Flag Value", values = flag_colours, drop = FALSE) +
      facet_wrap(~depth, ncol = 1) +
      theme_light() +
      theme(
        legend.position = "none",
        strip.text = element_text(colour = "black", size = 10),
        strip.background = element_rect(fill = "white", colour = "darkgrey")
      ) #+
      #guides(color = guide_legend(override.aes = list(size = 4)))
    
    subchunkify(p, fig_height = h, fig_width = 8.5)
    
  }
}
```
:::

## Deployments of Interest

These figures show the flagged observations for each threshold for several deployments with notable dissolved oxygen signals.

::: panel-tabset

```{r}
#| results: "asis"

depl <- data.frame(
  station = c("Birchy Head", "Birchy Head", 
              "Flat Island", "Flat Island", 
              "Tilley Point"),
  depl_date = c("2020-Nov-08", "2021-Nov-26", 
                "2021-Nov-21", "2022-May-24",
                "2021-May-21")
  )


for(i in 1:nrow(depl)){
  depl_i <- depl[i, ]
  
  cat('\n###', paste(depl_i$station, depl_i$depl_date), '\n')
  
  for(j in seq_along(thresh_plot)) {
    
    cat('\n####', thresh_plot[j], '\n')
    
    dat_j <-  dat_summary %>% 
      filter(
        station == depl_i$station, 
        depl_date == depl_i$depl_date, 
        threshold == thresh_plot[j])  %>%
      mutate(
        depth = paste0(
          sensor_depth_at_low_tide_m, " m", 
          " (", sensor_serial_number, ")"),
        depth = ordered(depth, levels = gtools::mixedsort(unique(depth)))
      ) %>% 
      rename(sd_rolling_flag_value = threshold_value) %>% 
      qc_assign_flag_labels() 
    
    if(nrow(dat_j) > 0) {
      
      # height of figure j
      n_sensor <- length(unique(dat_j$sensor_serial_number))
      
      if(n_sensor == 1) h = 2.75
      if(n_sensor > 1) h = 2.75 * n_sensor
      
      p <- ggplot(
        dat_j, 
        aes(timestamp_utc, value_dissolved_oxygen_uncorrected_mg_per_l, 
            colour = sd_rolling_flag_value )
      ) +
        geom_point()+
        scale_y_continuous("Dissolved Oxygen (mg/L)") +
        scale_x_datetime("Date") +
        scale_colour_manual("Flag Value", values = flag_colours, drop = FALSE) +
        facet_wrap(~depth, ncol = 1) +
        theme_light() +
        theme(
          legend.position = "none",
          strip.text = element_text(colour = "black", size = 10),
          strip.background = element_rect(fill = "white", colour = "darkgrey")
        ) 
    
       subchunkify(p, fig_height = h, fig_width = 8.5)
    }
  }
}
```
:::

## Conclusions

- As with the dissolved oxygen - percent saturation, q_99.7 does a good job identifying the most obvious biofouling signals
  - See deployments of interest.
  - However, this threshold leaves gaps of "Good" observations between consistent "Suspect.
  - This threshold misses some suspected biofouling 
    - Birchy Head 2021-Nov-26, Flat Island 2021-Nov-21, Tilley Point 2021-May-21

- q_90 would be a good option, although it occasionally results in what are likely false positives.

- It is recommended to use q_95, which provides a reasonable balance between false positives and false negatives.
  - Same quantile used for dissolved oxygen - percent saturation data.
  - Flags approximately the same percent of observations as the selected dissolved oxygen - percent saturation threshold.


