---
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(DT)
library(here)
library(ggplot2)
library(leaflet)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(readr)
library(summaryplots)

source(here("functions/subchunkify.R"))

theme_set(theme_light())

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

# summarized data 
dat <- read_csv(
  here("data/4_summary_filtered_data_do_additional_qc.csv"), 
  show_col_types = FALSE
) %>% 
  filter(variable == "Dissolved Oxygen", units == "percent saturation")

# station locations
st_locations <- read_csv(
  here("data/Station_Locations_2022-12-06.csv"), show_col_types = FALSE
)

county_pal <- get_county_colour_palette(length(unique(st_locations$COUNTY)))

theme_facet_plotly <- theme(panel.spacing.y = unit(15, "lines"))
```

# EXPLORATORY ANALYSIS: DISSOLVED OXYGEN 

*Strict QC*

July 25, 2023

- Data submitted to the Open Data Portal in December 2022.
- Depth was rounded to nearest whole number
  - Results in depth of 0 m for 1042 and Sissiboo.
- Preliminary QC was applied to the data.
  - Obvious outliers removed.
  - Suspected biofouling removed.
- Additional QC applied was based on review of 1.1_explore_do_data.html
  - Fresh water stations not considered.
    - Piper Lake, Hourglass Lake, Sissiboo
- *Additional QC applied was based on distributions in 3.0_do_rolling_sd.html.
  - Removed counties with limited deployments.
    - Antigonish, Colchester, Inverness, Pictou, Queens
  - Removed Digby because additional review suggests substantial suspicious observations.*
  
# Station Locations

Approximate location of stations with dissolved oxygen data.
Marker size is proportional to the number of dissolved oxygen observations within the county.

```{r}
#| fig-height: 7

# join the station locations dataset with the number of obs from each station
st_locations <- st_locations %>% 
  inner_join(
    dat %>% 
      filter(group == "all_station") %>% 
      select(COUNTY = county, STATION = station, n),
    by = c("COUNTY", "STATION") 
  ) %>% 
  group_by(COUNTY) %>% 
  mutate(
    n_prop = round(n / sum(n), digits = 2),
    popup = paste(COUNTY, STATION, n_prop, sep = "</br>")
  ) 

# interactive map
leaflet(st_locations) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  add_circle_markers_county(
    county_pal = county_pal, 
    size =  ~n_prop * 25
  ) %>% 
  add_county_fill_legend(county_pal = county_pal)
```

- `r nrow(dat %>% filter(group == "all_station") %>% group_by(county) %>% filter(n() == 1))` counties have only 1 station with dissolved oxygen data: `r x <- dat %>% filter(group == "all_station") %>% group_by(county) %>% filter(n() == 1); paste(x$county, collapse = ", ")`
- Guysborough has `r nrow(dat %>% filter(group == "all_station", county == "Guysborough") %>% distinct(station))` stations with dissolved oxygen data.

# Gross Range

```{r}
county <- dat %>% 
  filter(group == "county" | group == "all_data") %>% 
  mutate(
    county = if_else(group == "all_data", "All", county),
    n_percent = round(n * 100 / max(n), digits = 1)
  ) %>% 
  select(-c(year, depth, month, station))
```

## County

::: panel-tabset

### Figure 1
```{r}
#| fig-height: 8
#| fig-cap: The number of dissolved oxygens observations in each county.
#| fig-cap-location: top

p <- county  %>% 
  filter(county != "All") %>% 
  plot_n_obs_county()

ggplotly(p)
```

### Figure 2
```{r}
#| fig-height: 8
#| fig-cap: The mean and standard deviation of dissolved oxygen in each county.
#| fig-cap-location: top

p <- plot_mean_sd_county(county)

ggplotly(p, tooltip = "text")
```

### Table 1
```{r}
#| fig-cap: Summary statistics of dissolved oxygen observations - total and by county. Grossrange user-defined thresholds are calculated as user_min = mean - 3 * stdev and user_max = mean + 3 * stdev.
#| fig-cap-location: top

county %>% 
  select(county, n, mean, stdev) %>% 
  mutate(user_min = mean - 3 * stdev, user_max = mean + 3 * stdev) %>% 
  datatable(options = dt_options, rownames = FALSE)
```
:::

- Narrower range of user thresholds for pooled data than filtered data (87.2,	121.3) and shallow data (87.3, 121.5).

## Depth

```{r}
depth <- dat %>% 
  filter(group == "all_depth") %>% 
  mutate(n_percent = round(n * 100 / sum(n), digits = 1)) %>% 
  select(-c(month, group, station, units, county))
```

::: panel-tabset

### Figure 3
```{r}
#| fig-height: 10
#| fig-cap: Number of dissolved oxygen observations at each depth.
#| fig-cap-location: top

p <- plot_n_obs_depth(depth, facet_county = FALSE)

ggplotly(p, tooltip = "text")
```

### Figure 4
```{r}
#| fig-height: 10
#| fig-cap: Mean and standard deviation of dissolved oxygen at each depth.
#| fig-cap-location: top

p <- plot_mean_sd_depth(depth, facet_county = FALSE, text_size = 12)

ggplotly(p, tooltip = "text")
```
:::

- Dissolved oxygen sensors are typically placed at 5 m below the surface.


## County + Depth

```{r}
county_depth <- dat %>% filter(group == "county_depth") 
```

::: panel-tabset

### Figure 5
```{r}
#| fig.height: 8
#| fig-cap: Depths with dissolved oxygen measurements by county.
#| fig-cap-location: top

p <- ggplot(county_depth, aes(depth, county, col = factor(depth))) +
  geom_point(size = 4) +
  scale_y_discrete(name = "", limits = rev) +
  scale_x_continuous("Sensor Depth Below the Surface at Low Tide (m)")

ggplotly(p)
```

### Figure 6
```{r}
#| fig-height: 10
#| fig-cap: The number of dissolved oxygen obervations at each depth for each county.
#| fig-cap-location: top

p <- plot_n_obs_depth(county_depth) +
  theme_facet_plotly 

ggplotly(p)
```

### Figure 7
```{r}
#| fig-height: 10
#| fig-cap: Average dissolved oxygen and standard deviation for each county at each measured depth. 

p <- plot_mean_sd_depth(county_depth, text_size = 12) +
  theme_facet_plotly 

ggplotly(p, tooltip = "text")
```

### Figure 8

```{r}
#| fig-height: 8
#| fig-cap: Average dissolved oxygen and standard deviation of all observations (grey) and average dissolved oxygen at each depth.

p <- ggplot(filter(county, county != "All"), aes(mean, county)) +
  geom_errorbar(
    aes(xmin = mean - stdev, xmax = mean + stdev), 
    width = 0, linewidth = 1, col = "grey70"
  ) +
  geom_point(size = 5, pch = 21, fill = "grey70", col = "grey50") +
  geom_point(
    data = county_depth, 
    aes(mean, county, col = factor(depth)), 
    alpha = 0.75, size = 2
  ) +
  scale_y_discrete(name = "", limits = rev) +
  scale_x_continuous("Dissolved Oxygen (percent saturation)")

ggplotly(p)
```
:::

# Climatology

## Month
```{r}
month <- dat %>% 
  filter(group == "all_month") %>% 
  mutate(
    n_percent = round(n * 100 / sum(n), digits = 1),
    month = month(month, label = TRUE)
  ) 
```

::: panel-tabset

### Figure 9
```{r}
#| fig-height: 8
#| fig-cap: The number of dissolved oxygens observations in each month. The number at the end of each bar indicates the percent of the total dissolved oxygen observations from the associated month.
#| fig-cap-location: top

p <- plot_n_obs_month(month, facet_county = FALSE)

ggplotly(p)
```

### Figure 10
```{r}
#| fig-height: 8
#| fig-cap-location: top

p <- plot_mean_sd_season(month, facet_county = FALSE)

ggplotly(p)
```

### Table 2
```{r}
#| fig-cap: Summary statistics of dissolved oxygen observations.
#| fig-cap-location: top

month %>% 
  select(month, n, mean, stdev) %>% 
  mutate(season_min = mean - 3 * stdev, season_max = mean + 3 * stdev) %>% 
  datatable(options = dt_options, rownames = FALSE)
```

:::

- Fewest dissolved oxygen observations in the spring and summer, likely due to preliminary QC to remove biofouling.


## Month + Year 

```{r}
month_year <- dat %>% 
  filter(group == "all_month_year") %>%
  group_by(year) %>% 
  mutate(
    n_percent = round(n * 100 / sum(n), digits = 1),
    month = month(month, label = TRUE)
  ) 
```

::: panel-tabset

### Figure 11
```{r}
#| fig-height: 10
#| fig-cap: The number of dissolved oxygen observations for each month by year.
#| fig-cap-location: top

p <- plot_n_obs_month(month_year, text_size = 12) +
  facet_wrap(~year) +
  theme(panel.spacing.y = unit(15, "lines"))

ggplotly(p)
```

### Figure 12
```{r}
#| fig-height: 10
#| fig-cap: Monthly climatology by year.
#| fig-cap-location: top

p <- plot_mean_sd_season(month_year) +
  facet_wrap(~year) +
  theme(panel.spacing.y = unit(15, "lines"))

ggplotly(p)
```

### Figure 13
```{r}
#| fig-height: 8
#| fig-cap: Monthly climatology by year - interactive.
#| fig-cap-location: top

month_year <- month_year %>% 
  mutate(year = factor(year))

p <- plot_mean_sd_season_ribbon(month_year, group = "year", facet_county = FALSE)

ggplotly(p, tooltip = "text")
```

### Table 3

```{r}
month_year %>% 
  select(year, month, n, mean, stdev) %>% 
  datatable(options = dt_options, rownames = FALSE)
```
:::

- There is a small but noticeable climatological cycle.
- Difficult to parse whether this is a true climatological cycle, or a function of not trimming all biofouling.

## Month + County
```{r}
county_month <- dat %>% 
  filter(group == "county_month") %>% 
  mutate(month = month(month, label = TRUE)) %>% 
  select(-c(units, group, year, station, depth))
```

::: panel-tabset

### Figure 14
```{r}
#| fig-height: 12
#| fig-cap: The number of dissolved oxygen observations in each month for each county.
#| fig-cap-location: top

p <- plot_n_obs_month(county_month, text_size = 12) +
  theme(panel.spacing.y = unit(15, "lines"))

ggplotly(p)
```

### Figure 15
```{r}
#| fig-height: 10
#| fig-cap: The mean and standard deviation of dissolved oxygen in each county.
#| fig-cap-location: top

p <- plot_mean_sd_season(county_month) +
  theme(panel.spacing.y = unit(15, "lines"))

ggplotly(p)
```

### Figure 16
```{r}
#| fig-height: 8
#| fig-cap: The mean and standard deviation of dissolved oxygen in each county.
#| fig-cap-location: top

p <- plot_mean_sd_season_ribbon(
  county_month, facet_county = FALSE, group = "county", text_size = 12
)

ggplotly(p, tooltip = "text")
```

### Figure 17
```{r}
#| fig-height: 8
#| fig-cap: The mean and standard deviation of dissolved oxygen in each county.
#| fig-cap-location: top

county_gr <- county_month %>% 
  rename(mean_month = mean) %>% 
  group_by(county) %>%
  summarise(
    mean = round(mean(mean_month), digits = 3),
    stdev = round(sd(mean_month), digits = 3),
    group = "county_climatology"
  ) %>% 
  bind_rows(
    county %>% 
      filter(group != "all_data") %>% 
      select(-c(n, n_percent))
  ) 

p <- ggplot(county_gr, aes(mean, county, fill = group, col = group)) +
  geom_errorbar(
    aes(xmin = mean - stdev, xmax = mean + stdev), 
    width = 0, linewidth = 1, alpha = 0.75
  ) +
  geom_point(size = 3, pch = 21, col = 1) +
  scale_y_discrete(name = "", limits = rev) +
  scale_x_continuous("Dissolved Oxygen (percent saturation)")

ggplotly(p)
```

### Figure 18
```{r}
#| fig-height: 8
#| fig-cap: The mean and standard deviation of dissolved oxygen in each county.
#| fig-cap-location: top

p <- ggplot(county_gr, aes(stdev, county, fill = group)) +
  geom_col(position = "dodge") +
  scale_y_discrete(name = "", limits = rev) +
  scale_x_continuous("Standard Deviation (percent saturation)")

ggplotly(p)
```
:::

- Fewest dissolved oxygen observations in the spring and summer, likely due to biofouling.
- Small but noticeable climatological cycle within counties.
- Very small standard deviations
- Some indication that the number of observations per month impacts the mean.
- 1 county missing at least 1 month of data (Annapolis)


## Month + County + Year
```{r}
county_month_year <- dat %>% 
  filter(group == "county_month_year") %>% 
  group_by(year, month) %>% 
  mutate(
    n_percent = round(n * 100 / sum(n), digits = 1),
    month = month(month, label = TRUE),
    year = factor(year)
  ) %>% 
  select(-c(units, group, station, depth)) %>% 
  ungroup()
```

### Number of Obervations

Note: x-axis is not consistent between figures.

Figures are not interactive.

::: panel-tabset

```{r}
#| results: "asis"

counties <- unique(county_month_year$county)

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- county_month_year %>% filter(county == county_i)
  
  # height of figure j
  n_years <- length(unique(dat_i$year))
  if(n_years <= 3) h = 4
  if(n_years > 3) h = 8
  
  cat('\n###', county_i, '\n')
  
  p <- plot_n_obs_month(dat_i, text_size = 12) +
    facet_wrap(~year, ncol = 3) 
  
  subchunkify(p, fig_height = h, fig_width = 8.5)
}
```
:::

### Climatology

Figures are not interactive.

::: panel-tabset

```{r}
#| fig-height: 6
#| results: "asis"

n_years <- length(unique(county_month_year$year))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  
  dat_i <- county_month_year %>% 
    filter(county == county_i)
  
  cat('\n###', county_i, '\n')
  
  p <- plot_mean_sd_season_ribbon(
    dat_i, group = "year",
    n_group = n_years,
    facet_county = FALSE, y_limits = c(60, 130)
  ) 
  
  print(p)
}
```
:::

