---
format: 
  html:
    toc: true
    toc-depth: 3
    embed-resources: true
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, dpi = 600, fig.width = 8)

library(dplyr)
library(DT)
library(here)
library(ggplot2)
library(leaflet)
library(lubridate)
library(plotly)
library(RColorBrewer)
library(readr)
library(strings)
library(summaryplots)
library(tidyr)
library(viridis)

source(here("functions/filter_out_suspect_obs.R"))
source(here("functions/remove_do_correction.R"))
source(here("functions/subchunkify.R"))

theme_set(theme_light())

x_axis <- scale_x_continuous("Salinity (PSU)")

dt_options <- list(
  dom = 'ft',
  paging = FALSE,
  searching = TRUE,
  scrollY = "500px",
  columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

# data - preliminary QC 
dat_all <- import_strings_data(input_path = here("data-raw")) %>%
  filter(VARIABLE == "Salinity") %>% 
  select(
    COUNTY,WATERBODY, STATION,  
    DEPLOYMENT_PERIOD, TIMESTAMP, DEPTH, 
    VARIABLE, VALUE, UNITS
  ) %>%
  mutate(
    DEPTH = factor(round(as.numeric(DEPTH))),
    MONTH = month(TIMESTAMP),
    YEAR = factor(year(TIMESTAMP))
  ) 

# filtered data
dat_filt <- dat_all %>% 
  filter_out_suspect_obs() 

# filtered data with no Inverness
dat_no_inverness <- dat_filt %>% filter(COUNTY != "Inverness")

# filtered data with ONLY Inverness
dat_inverness <- dat_filt %>% filter(COUNTY == "Inverness")


county_pal <- get_county_colour_palette(length(unique(dat_all$COUNTY)))
depth_pal <- viridis(length(unique(dat_all$DEPTH)), direction = -1)
year_pal <- viridis( length(unique(dat_all$YEAR)), option = "C")

theme_facet_plotly <- theme(
  panel.spacing.y = unit(30, "lines"),
  panel.spacing.x = unit(20, "lines")
)

```

# Salinity Distribution

August 30, 2023

- Data submitted to the Open Data Portal in December 2022.
- Preliminary QC (submitted to Open Data Portal)
  - Obvious outliers removed.
  - Suspected biofouling removed.
- Filtered Data
  - Additional suspect data trimmed (mostly drift over deployment)


## Preliminary QC 

### Pooled

::: panel-tabset

#### Figure 1: All observations
```{r}
#| label: fig-dist-all
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU).

p <- plot_histogram(dat_all, hist_col = VALUE, binwidth = 0.5) +
  x_axis

ggplotly(p)
```

#### Figure 2: By depth
```{r}
#| label: fig-dist-all-depth
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5,
  fill_col = DEPTH, pal = depth_pal
) +
  x_axis

ggplotly(p)
```

#### Figure 3: By year
```{r}
#| label: fig-dist-all-year
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, 
  fill_col = YEAR, pal = year_pal
) +
  x_axis

ggplotly(p)
```

:::

### County

::: panel-tabset

#### Figure 4: County 
```{r}
#| label: fig-dist-county
#| fig-height: 8
#| fig-cap: Distribution of salinity observations by county (binwidth = 0.5 PSU).

p <- dat_all %>% 
  plot_histogram(hist_col = VALUE, binwidth = 0.5) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

#### Figure 5: County + Depth
```{r}
#| label: fig-dist-county-depth
#| fig-height: 8
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), within each county, coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = DEPTH, pal = depth_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly
  
ggplotly(p)
```

#### Figure 6: County + Year
```{r}
#| label: fig-dist-county-year
#| fig-height: 8
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), within each county coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```
:::

#### County + Month + Year

::: panel-tabset
```{r}
#| fig-height: 10
#| results: "asis"

counties <- unique(dat_all$COUNTY)

dat_all <- dat_all %>% 
  mutate(MONTH = month(MONTH, label = TRUE))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- dat_all %>% 
    filter(COUNTY == county_i) 
  
  cat('\n###', county_i, '\n')
  
  p <- plot_histogram(
    dat_i, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
  ) +
    x_axis +
    facet_wrap(~MONTH, ncol = 3)  
  
  print(p)
}

```
:::

## Filtered Data

### Pooled

::: panel-tabset

#### Figure 7: All observations
```{r}
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU).

p <- plot_histogram(dat_filt, hist_col = VALUE, binwidth = 0.5) +
  x_axis

ggplotly(p)
```

#### Figure 8: By depth
```{r}
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), coloured by depth.

p <- plot_histogram(
  dat_filt, hist_col = VALUE, binwidth = 0.5,
  fill_col = DEPTH, pal = depth_pal
) +
  x_axis

ggplotly(p)
```

#### Figure 9: By year
```{r}
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, 
  fill_col = YEAR, pal = year_pal
) +
  x_axis

ggplotly(p)
```

:::

### County

::: panel-tabset

#### Figure 10: County 
```{r}
#| fig-height: 8
#| fig-cap: Distribution of salinity observations by county (binwidth = 0.5 PSU).

p <- dat_filt%>% 
  plot_histogram(hist_col = VALUE, binwidth = 0.5) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

#### Figure 11: County + Depth
```{r}
#| fig-height: 8
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), within each county, coloured by depth.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = DEPTH, pal = depth_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly
  
ggplotly(p)
```

#### Figure 12: County + Year
```{r}
#| fig-height: 8
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), within each county coloured by year.

p <- plot_histogram(
  dat_all, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```
:::

#### County + Month + Year

::: panel-tabset
```{r}
#| fig-height: 10
#| results: "asis"

counties <- unique(dat_all$COUNTY)

dat_filt<- dat_filt%>% 
  mutate(MONTH = month(MONTH, label = TRUE))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- dat_filt%>% 
    filter(COUNTY == county_i) 
  
  cat('\n###', county_i, '\n')
  
  p <- plot_histogram(
    dat_i, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
  ) +
    x_axis +
    facet_wrap(~MONTH, ncol = 3)  
  
  print(p)
}

```
:::

## Filtered Data - No Inverness

### Pooled

::: panel-tabset

#### Figure 13: All observations
```{r}
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU).

p <- plot_histogram(dat_no_inverness, hist_col = VALUE, binwidth = 0.5) +
  x_axis

ggplotly(p)
```

#### Figure 14: By depth
```{r}
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), coloured by depth.

p <- plot_histogram(
  dat_no_inverness, hist_col = VALUE, binwidth = 0.5,
  fill_col = DEPTH, pal = depth_pal
) +
  x_axis

ggplotly(p)
```

#### Figure 15: By year
```{r}
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), coloured by year.

p <- plot_histogram(
  dat_no_inverness, hist_col = VALUE, binwidth = 0.5, 
  fill_col = YEAR, pal = year_pal
) +
  x_axis

ggplotly(p)
```

:::

### County

::: panel-tabset

#### Figure 16: County 
```{r}
#| fig-height: 8
#| fig-cap: Distribution of salinity observations by county (binwidth = 0.5 PSU).

p <- dat_no_inverness %>% 
  plot_histogram(hist_col = VALUE, binwidth = 0.5) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```

#### Figure 17: County + Depth
```{r}
#| fig-height: 8
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), within each county, coloured by depth.

p <- plot_histogram(
  dat_no_inverness, hist_col = VALUE, binwidth = 0.5, fill_col = DEPTH, pal = depth_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly
  
ggplotly(p)
```

#### Figure 18: County + Year
```{r}
#| fig-height: 8
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU), within each county coloured by year.

p <- plot_histogram(
  dat_no_inverness, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
) +
  x_axis +
  facet_wrap(~COUNTY, scales = "free", ncol = 3) +
  theme_facet_plotly

ggplotly(p)
```
:::

#### County + Month + Year

::: panel-tabset
```{r}
#| fig-height: 10
#| results: "asis"

counties <- unique(dat_no_inverness$COUNTY)

dat_no_inverness <- dat_no_inverness %>% 
  mutate(MONTH = month(MONTH, label = TRUE))

for(i in seq_along(counties)){
  
  county_i <- counties[i]
  dat_i <- dat_no_inverness %>% 
    filter(COUNTY == county_i) 
  
  cat('\n###', county_i, '\n')
  
  p <- plot_histogram(
    dat_i, hist_col = VALUE, binwidth = 0.5, fill_col = YEAR, pal = year_pal
  ) +
    x_axis +
    facet_wrap(~MONTH, ncol = 3)  
  
  print(p)
}

```
:::

## Inverness

### Pooled

::: panel-tabset

#### Figure 19: All observations
```{r}
#| fig-height: 6
#| fig-cap: Distribution of salinity observations (binwidth = 0.5 PSU).

p <- plot_histogram(dat_inverness, hist_col = VALUE, binwidth = 0.5) +
  x_axis

ggplotly(p)
```

:::

# Recommendations

- Preliminary QC and Filtered Data pooled distributions are bimodal because of Inverness observations.



